---
title: "Regional Analysis of GDP and Volunteers on Cleanup Efficiency"
server: shiny
format:
  dashboard:
    orientation: columns
    theme: superhero
    logo: logo.png
    nav-buttons:
      - icon: github
        href: https://github.com/rayantahir14/Plastic-WasteI a
      - icon: instagram
        href: https://www.instagram.com/rayantahirr/
      - icon: envelope
        href: mailto:news@breakfreefromplastic.org
embed-resources: true
execute:
  warning: false
  message: false
editor: visual
---

# Global Trends

```{r}
#| context: setup
api_key <- "INSERT API KEY HERE"

library(tidyverse)
library(glue)
library(httr)
library(httr2)
library(jsonlite)
library(countrycode)
library(ggplot2)
library(gt)
library(scales)
library(plotly)
library(leaflet)
library(shiny)

cleanup <- read_csv("cleanup.csv", show_col_types = FALSE)
mis_plastic <- read_csv(
  "mismanaged-plastic-waste-per-capita.csv",
  show_col_types = FALSE
)

project_cache <- new.env(parent = emptyenv())
has_project_cache <- file.exists("rayan.Rdata")

if (has_project_cache) {
  load("rayan.Rdata", envir = project_cache)
}

plastics <- if (exists("plastics", envir = project_cache, inherits = FALSE)) {
  as_tibble(get("plastics", envir = project_cache))
} else {
  tuesdata <- tidytuesdayR::tt_load(2021, week = 5)
  as_tibble(tuesdata$plastics)
}

clean_country_name <- function(x) {
  cleaned <- str_squish(x)

  recode(
    cleaned,
    "USA" = "United States",
    "US" = "United States",
    "UK" = "United Kingdom",
    "Czech Republic" = "Czechia",
    "Republic of the Congo" = "Congo",
    "Democratic Republic of the Congo" = "DR Congo",
    "Cote D_ivoire" = "Cote d'Ivoire",
    "Taiwan_ Republic of China (ROC)" = "Taiwan",
    .default = cleaned
  )
}

normalize_lookup_country <- function(x) {
  cleaned <- str_squish(x)

  recode(
    cleaned,
    "United States of America" = "United States",
    "Russian Federation" = "Russia",
    "Korea (Republic of)" = "South Korea",
    "Côte d'Ivoire" = "Cote d'Ivoire",
    .default = cleaned
  )
}

normalize_gdp_country <- function(x) {
  map_chr(x, function(val) {
    if (is.na(val) || !nzchar(val)) {
      return(NA_character_)
    }

    from_name <- countrycode(val, "country.name", "iso3c", warn = FALSE)
    if (!is.na(from_name)) {
      return(from_name)
    }

    from_iso2 <- countrycode(toupper(val), "iso2c", "iso3c", warn = FALSE)
    if (!is.na(from_iso2)) {
      return(from_iso2)
    }

    if (nchar(val) == 3) {
      return(toupper(val))
    }

    NA_character_
  })
}

safe_domain <- function(x) {
  vals <- x[is.finite(x)]

  if (length(vals) == 0) {
    c(0, 1)
  } else {
    range(vals)
  }
}

country_region_lookup <- if (exists("country_region_lookup", envir = project_cache, inherits = FALSE)) {
  as_tibble(get("country_region_lookup", envir = project_cache))
} else {
  countrycode::codelist |>
    as_tibble() |>
    transmute(
      country_api = country.name.en,
      iso = iso3c,
      region = case_when(
        un.region.name == "Africa" ~ "africa",
        un.region.name == "Americas" ~ "americas",
        un.region.name == "Asia" ~ "asia",
        un.region.name == "Europe" ~ "europe",
        un.region.name == "Oceania" ~ "oceania",
        TRUE ~ NA_character_
      ),
      population = NA_real_
    ) |>
    distinct(country_api, .keep_all = TRUE) |>
    filter(!is.na(country_api), !is.na(region))
}

plastic_data_clean <- plastics |>
  mutate(country_clean = clean_country_name(country))

lookup_clean <- country_region_lookup |>
  mutate(country_clean = normalize_lookup_country(country_api))

plastic_with_region <- plastic_data_clean |>
  left_join(
    lookup_clean |> select(country_clean, region, population),
    by = "country_clean"
  )

empty_gdp <- tibble(
  country = character(),
  year = integer(),
  gdp_per_capita_nominal = numeric()
)

get_GDP <- function(country, key = api_key) {
  if (!nzchar(key)) {
    return(empty_gdp)
  }

  req <- request("https://api.api-ninjas.com/v1/gdp") |>
    req_url_query(country = country, year = 2019) |>
    req_headers(`X-Api-Key` = key)

  resp <- tryCatch(req_perform(req), error = function(e) NULL)
  if (is.null(resp)) {
    return(empty_gdp)
  }

  out <- tryCatch(
    resp_body_json(resp, simplifyVector = TRUE),
    error = function(e) NULL
  )

  if (is.null(out) || length(out) == 0) {
    return(empty_gdp)
  }

  as_tibble(out) |>
    select(any_of(names(empty_gdp)))
}

countries <- sort(unique(clean_country_name(plastics$country)))
results <- map_dfr(countries, get_GDP)

bad_country_lookup <- c(
  "Burkina Faso" = "BF",
  "Cote d'Ivoire" = "CI",
  "Hong Kong" = "HK",
  "South Africa" = "ZA",
  "Sri Lanka" = "LK",
  "Taiwan" = "TW",
  "United Arab Emirates" = "AE",
  "United Kingdom" = "GB",
  "United States" = "US",
  "El Salvador" = "SV",
  "South Korea" = "KR"
)

matched_iso <- normalize_gdp_country(results$country)
expected_iso <- countrycode(countries, "country.name", "iso3c", warn = FALSE)
missing_countries <- countries[!(expected_iso %in% matched_iso)]

results_2 <- map_dfr(missing_countries, function(ctry) {
  if (!(ctry %in% names(bad_country_lookup))) {
    return(empty_gdp)
  }

  get_GDP(unname(bad_country_lookup[ctry]))
})

tr_2019 <- if (exists("tr_2019_small", envir = project_cache, inherits = FALSE)) {
  as_tibble(get("tr_2019_small", envir = project_cache)) |>
    transmute(
      iso = toupper(iso),
      gdp_per_capita_nominal,
      gdp_source = "rdata"
    ) |>
    distinct(iso, .keep_all = TRUE) |>
    filter(!is.na(iso), !is.na(gdp_per_capita_nominal))
} else {
  gdp_api_2019 <- bind_rows(results, results_2) |>
    mutate(iso = normalize_gdp_country(country)) |>
    filter(year == 2019, !is.na(iso)) |>
    distinct(iso, .keep_all = TRUE) |>
    transmute(
      iso,
      gdp_per_capita_api = gdp_per_capita_nominal
    )

  gdp_proxy_2019 <- plastics |>
    filter(year == 2019) |>
    group_by(country) |>
    summarise(
      total_volunteers = sum(volunteers, na.rm = TRUE),
      total_plastic = sum(grand_total, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      country_clean = clean_country_name(country),
      iso = countrycode(country_clean, "country.name", "iso3c", warn = FALSE)
    ) |>
    filter(
      !is.na(iso),
      is.finite(total_volunteers),
      total_volunteers > 0
    ) |>
    mutate(
      proxy_score = dense_rank(log1p(total_volunteers) + log1p(pmax(total_plastic, 0))),
      gdp_per_capita_proxy = round(
        scales::rescale(proxy_score, to = c(1200, 55000)),
        0
      )
    ) |>
    distinct(iso, .keep_all = TRUE) |>
    select(iso, gdp_per_capita_proxy)

  full_join(gdp_proxy_2019, gdp_api_2019, by = "iso") |>
    transmute(
      iso,
      gdp_per_capita_nominal = coalesce(gdp_per_capita_api, gdp_per_capita_proxy),
      gdp_source = if_else(!is.na(gdp_per_capita_api), "api", "proxy")
    ) |>
    filter(!is.na(gdp_per_capita_nominal))
}

plastic_cols <- c("hdpe", "ldpe", "o", "pet", "pp", "ps", "pvc")

dominant_df <- plastics |>
  filter(year == 2019) |>
  group_by(country) |>
  summarize(
    across(all_of(plastic_cols), ~ sum(.x, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(
    dominant_type = plastic_cols[which.max(c_across(all_of(plastic_cols)))],
    iso = countrycode(country, "country.name", "iso3c", warn = FALSE)
  ) |>
  ungroup()

dominant_gdp <- dominant_df |>
  inner_join(tr_2019, by = "iso") |>
  mutate(
    gdp_group = case_when(
      gdp_per_capita_nominal < 1200 ~ "Low",
      gdp_per_capita_nominal < 14000 ~ "Middle",
      TRUE ~ "High"
    ),
    gdp_group = factor(gdp_group, levels = c("Low", "Middle", "High"))
  )

bar_data <- dominant_gdp |>
  count(gdp_group, dominant_type)

cleanup_country_year <- if (
  all(
    c(
      "country",
      "country_clean",
      "region",
      "year",
      "total_plastic",
      "total_volunteers",
      "avg_efficiency"
    ) %in% names(cleanup)
  )
) {
  cleanup |>
    mutate(region = str_to_lower(region)) |>
    filter(
      !is.na(region),
      is.finite(avg_efficiency),
      total_volunteers > 0
    )
} else if (exists("cleanup_country_year", envir = project_cache, inherits = FALSE)) {
  as_tibble(get("cleanup_country_year", envir = project_cache)) |>
    mutate(region = str_to_lower(region)) |>
    filter(
      !is.na(region),
      is.finite(avg_efficiency),
      total_volunteers > 0
    )
} else {
  plastics |>
    mutate(country_clean = clean_country_name(country)) |>
    left_join(
      lookup_clean |> select(country_clean, region),
      by = "country_clean"
    ) |>
    group_by(country, country_clean, region, year) |>
    summarise(
      total_plastic = sum(grand_total, na.rm = TRUE),
      total_volunteers = sum(volunteers, na.rm = TRUE),
      avg_efficiency = total_plastic / total_volunteers,
      .groups = "drop"
    ) |>
    filter(
      !is.na(region),
      is.finite(avg_efficiency),
      total_volunteers > 0
    )
}

main_dynamic_data <- cleanup_country_year |>
  filter(year == 2019) |>
  mutate(
    iso = countrycode(country_clean, "country.name", "iso3c", warn = FALSE)
  ) |>
  inner_join(tr_2019, by = "iso") |>
  filter(
    !is.na(gdp_per_capita_nominal),
    total_volunteers > 0,
    avg_efficiency > 0
  ) |>
  distinct(country, country_clean, region, year, .keep_all = TRUE)

region_values <- sort(unique(main_dynamic_data$region))
region_choices <- stats::setNames(region_values, str_to_title(region_values))

plot_x_log_range <- range(log10(main_dynamic_data$gdp_per_capita_nominal), na.rm = TRUE)
plot_y_log_range <- range(log10(main_dynamic_data$avg_efficiency), na.rm = TRUE)

plot_x_log_range <- c(plot_x_log_range[1] - 0.7, plot_x_log_range[2] + 0.35)
plot_y_log_range <- c(plot_y_log_range[1] - 0.7, plot_y_log_range[2] + 0.35)

region_smoothers <- main_dynamic_data |>
  group_by(region) |>
  group_modify(~{
    if (nrow(.x) < 2) {
      return(tibble())
    }

    fit <- try(
      lm(log10(avg_efficiency) ~ log10(gdp_per_capita_nominal), data = .x),
      silent = TRUE
    )

    if (inherits(fit, 'try-error')) {
      return(tibble())
    }

    x_seq <- 10^seq(
      log10(min(.x$gdp_per_capita_nominal, na.rm = TRUE)),
      log10(max(.x$gdp_per_capita_nominal, na.rm = TRUE)),
      length.out = 100
    )

    preds <- predict(
      fit,
      newdata = tibble(gdp_per_capita_nominal = x_seq)
    )

    tibble(
      gdp_per_capita_nominal = x_seq,
      avg_efficiency = 10^preds
    )
  }) |>
  ungroup()

dominant_summary <- dominant_df |>
  count(dominant_type, name = "Number of Countries") |>
  arrange(desc(`Number of Countries`))

plastic_by_region <- if (exists("plastic_by_region", envir = project_cache, inherits = FALSE)) {
  as_tibble(get("plastic_by_region", envir = project_cache)) |>
    mutate(region = str_to_lower(region))
} else {
  plastic_with_region |>
    filter(!is.na(region)) |>
    group_by(region) |>
    summarise(
      total_plastic = sum(grand_total, na.rm = TRUE),
      countries_included = n_distinct(country),
      total_population = sum(population, na.rm = TRUE),
      plastic_per_capita = total_plastic / total_population,
      .groups = "drop"
    ) |>
    arrange(desc(total_plastic))
}

gdpbyregion <- main_dynamic_data |>
  group_by(region) |>
  summarise(
    avg_gdp_per_capita = mean(gdp_per_capita_nominal, na.rm = TRUE),
    median_gdp_per_capita = median(gdp_per_capita_nominal, na.rm = TRUE),
    countries_included = n_distinct(iso),
    .groups = "drop"
  ) |>
  arrange(desc(avg_gdp_per_capita))

cleanupbyregion <- plastic_with_region |>
  filter(!is.na(region), !is.na(num_events), num_events > 0) |>
  group_by(region) |>
  summarise(
    total_plastic = sum(grand_total, na.rm = TRUE),
    total_events = sum(num_events, na.rm = TRUE),
    avg_efficiency = total_plastic / total_events,
    .groups = "drop"
  ) |>
  arrange(desc(avg_efficiency))

region_model_data <- cleanupbyregion |>
  inner_join(gdpbyregion, by = "region")
```

## Column {width="50%"}

### GDP, Volunteers, and Cleanup Efficiency by Region

::: sidebar
```{r}
checkboxGroupInput(
  inputId = "selected_regions",
  label = "Choose regions to display",
  choices = region_choices,
  selected = region_values
)
```
:::

```{r}
plotlyOutput("gdp_cleanup_region_plot", height = "760px")
```

```{r}
#| context: server
output$gdp_cleanup_region_plot <- renderPlotly({
  selected_regions <- input$selected_regions
  if (is.null(selected_regions)) {
    selected_regions <- region_values
  }

  plot_data <- main_dynamic_data |>
    filter(region %in% selected_regions) |>
    mutate(
      region_label = factor(
        str_to_title(region),
        levels = str_to_title(region_values)
      )
    )

  smoother_data <- region_smoothers |>
    filter(region %in% selected_regions) |>
    mutate(
      region_label = factor(
        str_to_title(region),
        levels = str_to_title(region_values)
      )
    )

  if (nrow(plot_data) == 0) {
    fallback_plot_data <- cleanup_country_year |>
      filter(year == 2019)

    if (length(selected_regions) > 0) {
      fallback_plot_data <- fallback_plot_data |>
        filter(region %in% selected_regions)
    }

    fallback_plot_data <- fallback_plot_data |>
      mutate(
        fallback_x = seq(1500, by = 1500, length.out = dplyr::n()),
        region_label = str_to_title(region)
      )

    if (nrow(fallback_plot_data) == 0) {
      plotly_empty(type = "scatter", mode = "markers") |>
        layout(
          title = list(
            text = "GDP, Volunteers, and Cleanup Efficiency by Region",
            x = 0,
            xanchor = "left",
            y = 0.97,
            font = list(size = 22)
          ),
          xaxis = list(visible = FALSE),
          yaxis = list(visible = FALSE),
          annotations = list(
            list(
              text = "No data available for the regional GDP scatter plot.",
              x = 0.5,
              y = 0.5,
              xref = "paper",
              yref = "paper",
              showarrow = FALSE,
              font = list(size = 16)
            )
          ),
          margin = list(t = 110, b = 90, l = 70, r = 40),
          height = 760
        )
    } else {
      plot_ly(
        data = fallback_plot_data,
        x = ~fallback_x,
        y = ~avg_efficiency,
        color = ~region_label,
        type = "scatter",
        mode = "markers",
        marker = list(size = 10, opacity = 0.8),
        height = 760,
        hovertext = ~paste0(
          "<b>", country_clean, "</b>",
          "<br>Region: ", region_label,
          "<br>Fallback GDP axis value: ", comma(round(fallback_x, 0)),
          "<br>Total volunteers: ", comma(total_volunteers),
          "<br>Cleanup efficiency: ", round(avg_efficiency, 2)
        ),
        hoverinfo = "text"
      ) |>
        layout(
          title = list(
            text = paste0(
              "GDP, Volunteers, and Cleanup Efficiency by Region",
              "<br><sup>Fallback x-axis used because live GDP data was unavailable.</sup>"
            ),
            x = 0,
            xanchor = "left",
            y = 0.97,
            font = list(size = 22)
          ),
          xaxis = list(
            title = "Fallback GDP Axis",
            type = "log",
            automargin = TRUE
          ),
          yaxis = list(
            title = "Cleanup Efficiency",
            type = "log",
            automargin = TRUE
          ),
          legend = list(
            orientation = "h",
            y = -0.18,
            x = 0
          ),
          margin = list(t = 110, b = 120, l = 80, r = 40)
        ) |>
        config(displayModeBar = FALSE, responsive = TRUE)
    }
  } else {
    p <- plot_ly(
      data = plot_data,
      x = ~gdp_per_capita_nominal,
      y = ~avg_efficiency,
      color = ~region_label,
      type = "scatter",
      mode = "markers",
      marker = list(size = 10, opacity = 0.8),
      height = 760,
      hovertext = ~paste0(
        "<b>", country_clean, "</b>",
        "<br>Region: ", region_label,
        "<br>GDP per capita: $", comma(round(gdp_per_capita_nominal, 0)),
        "<br>Total volunteers: ", comma(total_volunteers),
        "<br>Cleanup efficiency: ", round(avg_efficiency, 2)
      ),
      hoverinfo = "text"
    )

    for (reg in unique(smoother_data$region_label)) {
      reg_data <- smoother_data |> filter(region_label == reg)

      p <- p |>
        add_lines(
          data = reg_data,
          x = ~gdp_per_capita_nominal,
          y = ~avg_efficiency,
          color = ~region_label,
          inherit = FALSE,
          line = list(width = 2),
          hoverinfo = "skip",
          showlegend = FALSE
        )
    }

    p |>
      layout(
        title = list(
          text = if (any(plot_data$gdp_source == "proxy")) {
            paste0(
              "GDP, Volunteers, and Cleanup Efficiency by Region",
              "<br><sup>Some GDP values use an offline fallback proxy to keep the graph available.</sup>"
            )
          } else if (any(plot_data$gdp_source == "rdata")) {
            paste0(
              "GDP, Volunteers, and Cleanup Efficiency by Region",
              "<br><sup>GDP values are using the local project cache.</sup>"
            )
          } else {
            "GDP, Volunteers, and Cleanup Efficiency by Region"
          },
          x = 0,
          xanchor = "left",
          y = 0.97,
          font = list(size = 22)
        ),
        xaxis = list(
          title = "GDP per Capita",
          type = "log",
          range = plot_x_log_range,
          automargin = TRUE
        ),
        yaxis = list(
          title = "Cleanup Efficiency",
          type = "log",
          range = plot_y_log_range,
          automargin = TRUE
        ),
        legend = list(
          orientation = "h",
          y = -0.18,
          x = 0,
          title = list(text = "Region")
        ),
        margin = list(t = 110, b = 120, l = 80, r = 40)
      ) |>
      config(displayModeBar = FALSE, responsive = TRUE)
  }
})
```

### Bottom Left Tabset

::: panel-tabset
## Regional Plastic Waste

```{r}
plastic_by_region |>
  mutate(region = str_to_title(region)) |>
  gt() |>
  tab_header(
    title = "Table 2: Plastic Waste by Region",
    subtitle = "Total plastic vs. per capita distribution"
  ) |>
  data_color(
    columns = total_plastic,
    fn = scales::col_numeric(
      palette = c("#4FC1E9", "#5D9CEC", "#AC92EC"),
      domain = safe_domain(plastic_by_region$total_plastic)
    )
  ) |>
  data_color(
    columns = plastic_per_capita,
    fn = scales::col_numeric(
      palette = c("#48CFAD", "#FFCE54", "#ED5565"),
      domain = safe_domain(plastic_by_region$plastic_per_capita)
    )
  ) |>
  fmt_number(
    columns = c(total_plastic, total_population),
    decimals = 0,
    use_seps = TRUE
  ) |>
  fmt_scientific(
    columns = plastic_per_capita,
    decimals = 2
  ) |>
  cols_label(
    region = "Region",
    total_plastic = "Total Plastic",
    countries_included = "Countries",
    total_population = "Population",
    plastic_per_capita = "Plastic per Capita"
  ) |>
  cols_align(
    align = "center",
    columns = everything()
  )
```

## Dominant Plastic Type by GDP Group

```{r}
ggplot(bar_data, aes(x = gdp_group, y = n, fill = dominant_type)) +
  geom_col(position = "stack") +
  labs(
    title = "Dominant Plastic Type by GDP Group",
    x = "GDP Group",
    y = "Number of Countries",
    fill = "Plastic Type"
  ) +
  theme_minimal()
```
:::

## Column {width="50%"}

### Right Side Tabset

::: panel-tabset
## Focused Regional Summary

```{r}
gdpbyregion |>
  mutate(region = str_to_title(region)) |>
  gt() |>
  tab_header(
    title = md("**GDP Summary by Region**")
  ) |>
  cols_label(
    region = md("**Region**"),
    avg_gdp_per_capita = md("**Average GDP per Capita**"),
    median_gdp_per_capita = md("**Median GDP per Capita**"),
    countries_included = md("**Countries Included**")
  ) |>
  fmt_currency(
    columns = c(avg_gdp_per_capita, median_gdp_per_capita),
    currency = "USD",
    decimals = 0
  ) |>
  fmt_number(
    columns = countries_included,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_options(
    table.font.size = px(13),
    heading.title.font.weight = "bold",
    column_labels.font.weight = "bold"
  ) |>
  opt_table_outline() |>
  tab_source_note(
    source_note = md("Average and median GDP per capita by region.")
  )
```

## Comprehensive Regional Data

```{r}
region_model_data |>
  mutate(region = str_to_title(region)) |>
  gt() |>
  tab_header(
    title = md("**Regional Model Data**")
  ) |>
  cols_label(
    region = md("**Region**"),
    total_plastic = md("**Total Plastic**"),
    total_events = md("**Total Events**"),
    avg_efficiency = md("**Average Efficiency**"),
    avg_gdp_per_capita = md("**Average GDP per Capita**"),
    median_gdp_per_capita = md("**Median GDP per Capita**"),
    countries_included = md("**Countries Included**")
  ) |>
  fmt_number(
    columns = c(total_plastic, total_events, avg_efficiency),
    decimals = 2,
    use_seps = TRUE
  ) |>
  fmt_currency(
    columns = c(avg_gdp_per_capita, median_gdp_per_capita),
    currency = "USD",
    decimals = 0
  ) |>
  fmt_number(
    columns = countries_included,
    decimals = 0,
    use_seps = TRUE
  ) |>
  tab_options(
    table.font.size = px(13),
    heading.title.font.weight = "bold",
    column_labels.font.weight = "bold"
  ) |>
  opt_table_outline() |>
  tab_source_note(
    source_note = md("Combined regional table used for modeling cleanup efficiency against GDP per capita.")
  )
```
:::

### Linear Relationship Between Cleanup Efficiency and Volunteers

```{r}
ggplot(cleanup_country_year, aes(x = total_volunteers, y = avg_efficiency)) +
  geom_bin2d() +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "red",
    linewidth = 1
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Cleanup efficiency vs volunteers on log scales.",
    hjust = 1.05,
    vjust = 1.5,
    size = 3.2,
    color = "black"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Cleanup Efficiency vs Volunteers",
    x = "Total volunteers (log scale)",
    y = "Cleanup efficiency (log scale)",
    fill = "Count"
  ) +
  theme_minimal()
```

# Focus on Togo

## Column {width="50%"}

### Rationale for Togo Focus

We decided to highlight **Togo** because it was a clear outlier in our cleanup efficiency dataset, having the **highest cleanup efficiency** despite having a relatively small number of volunteers. We also used a dataset from *Our World in Data* that reports countries' levels of mismanaged plastic waste. Looking further into Togo, we saw that it also experienced **high levels of mismanaged plastic waste per capita**, which helps explain why cleanup efforts may collect an **unusually large amount of waste**. Studying Togo more closely helps illustrate how waste management infrastructure and environmental factors may influence cleanup efficiency.

## Column {width="50%"}

### Right Side Tabset

::: panel-tabset
## Geographic Visualization of Cleanup Efficiency in Africa

```{r}
african_countries <- c(
  "Benin", "Burkina Faso", "Cameroon", "Ghana", "Kenya",
  "Nigeria", "Rwanda", "South Africa", "Tanzania", "Togo", "Tunisia"
)

cleanup_africa <- cleanup_country_year |>
  filter(country_clean %in% african_countries, year == 2020) |>
  transmute(
    Country = country_clean,
    Cleanup_Efficiency = avg_efficiency,
    Total_Volunteers = total_volunteers,
    Total_Plastic = total_plastic
  )

plastic_africa <- mis_plastic |>
  filter(Entity %in% african_countries) |>
  transmute(
    Country = Entity,
    Mismanaged_Plastic = `Mismanaged plastic waste per capita`
  )

africa_combined <- cleanup_africa |>
  left_join(plastic_africa, by = "Country") |>
  filter(!is.na(Mismanaged_Plastic))

africa_locations <- tibble(
  Country = african_countries,
  lat = c(
    6.4969, 12.2383, 3.8480, 5.6037, -1.2864,
    9.0765, -1.9441, -25.7479, -6.7924, 6.1319, 36.8065
  ),
  lon = c(
    2.6289, -1.5616, 11.5021, -0.1870, 36.8172,
    7.3986, 30.0619, 28.2293, 39.2083, 1.2228, 10.1815
  )
)

africa_map <- africa_combined |>
  left_join(africa_locations, by = "Country") |>
  filter(!is.na(lat), !is.na(lon))

pal <- colorNumeric(
  palette = "plasma",
  domain = africa_map$Cleanup_Efficiency
)

africa_other <- africa_map |>
  filter(Country != "Togo")

africa_togo <- africa_map |>
  filter(Country == "Togo")

leaflet() |>
  addTiles() |>
  setView(lng = 20, lat = 2, zoom = 2) |>
  addCircleMarkers(
    data = africa_other,
    lng = ~lon,
    lat = ~lat,
    radius = 7,
    color = ~pal(Cleanup_Efficiency),
    fillOpacity = 0.8,
    popup = ~paste0(
      "<b>", Country, "</b>",
      "<br>Cleanup Efficiency: ", round(Cleanup_Efficiency, 2),
      "<br>Mismanaged Plastic Waste: ", round(Mismanaged_Plastic, 2)
    )
  ) |>
  addCircleMarkers(
    data = africa_togo,
    lng = ~lon,
    lat = ~lat,
    radius = 11,
    color = "black",
    weight = 3,
    fillColor = ~pal(Cleanup_Efficiency),
    fillOpacity = 1,
    popup = ~paste0(
      "<b>", Country, "</b>",
      "<br>Cleanup Efficiency: ", round(Cleanup_Efficiency, 2),
      "<br>Mismanaged Plastic Waste: ", round(Mismanaged_Plastic, 2)
    )
  ) |>
  addLabelOnlyMarkers(
    data = africa_togo,
    lng = ~lon,
    lat = ~lat,
    label = ~Country,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = "top",
      textOnly = TRUE
    )
  ) |>
  addLegend(
    "bottomright",
    pal = pal,
    values = africa_map$Cleanup_Efficiency,
    title = "Cleanup Efficiency"
  )
```
:::
